\documentclass[nols,twoside,nofonts,nobib,nohyper]{tufte-handout}

\usepackage{fixltx2e}
\usepackage{tikz-cd}
\usepackage{tcolorbox}
\usepackage{appendix}
\usepackage{listings}
\lstset{language=TeX,
       frame=single,
       basicstyle=\ttfamily,
       captionpos=b,
       tabsize=4,
  }

\input{acronyms}
\renewcommand*{\acsfont}[1]{\textsc{#1}}

\usepackage[font=footnotesize]{caption}

\makeatletter
% Paragraph indentation and separation for normal text
\renewcommand{\@tufte@reset@par}{%
  \setlength{\RaggedRightParindent}{0pt}%
  \setlength{\JustifyingParindent}{0pt}%
  \setlength{\parindent}{0pt}%
  \setlength{\parskip}{\baselineskip}%
}
\@tufte@reset@par

% Paragraph indentation and separation for marginal text
\renewcommand{\@tufte@margin@par}{%
  \setlength{\RaggedRightParindent}{0pt}%
  \setlength{\JustifyingParindent}{0pt}%
  \setlength{\parindent}{0pt}%
  \setlength{\parskip}{\baselineskip}%
}
\makeatother

\usepackage{multicol}

\setcounter{secnumdepth}{3}

\title{Crossover ii\thanks{24.979: Topics in
    semantics\\\noindent\textit{Getting high: Scope, projection, and evaluation order}}}

\author[Patrick D. Elliott and Martin Hackl]{Patrick~D. Elliott \& Martin Hackl}

\addbibresource[location=remote]{/home/patrl/repos/bibliography/elliott_mybib.bib}

\lingset{
  belowexskip=0pt,
  aboveglftskip=0pt,
  belowglpreambleskip=0pt,
  belowpreambleskip=0pt,
  interpartskip=0pt,
  extraglskip=0pt,
  Everyex={\parskip=0pt}
}

\usepackage{float}


% \usepackage{booktabs} % book-quality tables
% \usepackage{units}    % non-stacked fractions and better unit spacing
% \usepackage{lipsum}   % filler text
% \usepackage{fancyvrb} % extended verbatim environments
%   \fvset{fontsize=\normalsize}% default font size for fancy-verbatim environments

% % Standardize command font styles and environments
% \newcommand{\doccmd}[1]{\texttt{\textbackslash#1}}% command name -- adds backslash automatically
% \newcommand{\docopt}[1]{\ensuremath{\langle}\textrm{\textit{#1}}\ensuremath{\rangle}}% optional command argument
% \newcommand{\docarg}[1]{\textrm{\textit{#1}}}% (required) command argument
% \newcommand{\docenv}[1]{\textsf{#1}}% environment name
% \newcommand{\docpkg}[1]{\texttt{#1}}% package name
% \newcommand{\doccls}[1]{\texttt{#1}}% document class name
% \newcommand{\docclsopt}[1]{\texttt{#1}}% document class option name
% \newenvironment{docspec}{\begin{quote}\noindent}{\end{quote}}% command specification environment

\begin{document}

\maketitle% this prints the handout title, author, and date

\begin{tcolorbox}{\textbf{Tentative revised schedule}}
  \tcblower

  We'll make an announcement on Friday about how the class will look for the
  rest of the semester, but the tentative plan is to spend a little bit more
  time talking about crossover, before moving on to presupposition and Haddock's puzzle.

  N.b. we skipped class in the first week of March, as I was away at a
  conference. This still needs to be rescheduled -- we'll be in touch about
  possible dates.

\end{tcolorbox}

\begin{tcolorbox}
  \textbf{Projects}
  \tcblower
  \textbf{Registered students:} thanks for submitting your project proposals. If
  you haven't already done so, please get in touch with either me, Martin (or
  both of us!) to arrange a time to talk through your proposal.
\end{tcolorbox}

\section{Last time -- Barker \& Shan: \textsc{wco} as a reflex of evaluation order}

At the beginning of the semester, we learned about a theory of scope-taking with
a built-in left-to-right bias -- \textit{continuation semantics}.

Concretely, due to the way that the composition rule \ac{sfa} was defined,
evaluation of quantificational effects \textit{mirrors} linear
order.\sidenote[][-7\baselineskip]{To cash this out, we needed to say something concrete about the
  syntax-semantics interface -- concretely, we committed to the ideas that (a)
  the basic combinatoric operation \textsc{Merge} is asymmetric, and (b) the
  syntactic and semantic composition proceed in lockstep (direct
  compositionality; \citealt{jacobson1999,jacobson_compositional_2014}).}

\ex~
\acf{sfa} (def.)\\
$\semtower{f []}{x} \ml{S} \semtower{g []}{y} ≔
\semtower{f [g []]}{x \ml{A} y}$
\xe

As we saw in the last class, an appealing consequence of this linear bias was a
natural account of \acf{wco} in terms of evaluation order.\sidenote{See
  especially \cite{shanBarker2006} and \cite[chapters 2 and 4]{barkerShan2015}.}

Recall, a simplified version of the \ac{wco} paradigm: scope can feed binding
(\ref{ex:wco1}), unless the binder doesn't precede the bound expression
(\ref{ex:wco2}).\sidenote{Throughout, i'll use superscript indices to indicate
  the binder, and subscript indices to indicate the bound expression(s).}

\ex
{}[Everyone$^{x}$'s mother] bought them$_{x}$ a bicycle.\label{ex:wco1}
\xe

\ex~
Their$_{x}$ mother showered everyone$^{x}$ with gifts.\\
\textit{cf. a different person showered everyone with gifts.}\label{ex:wco2}
\xe

The idea, briefly, was to generalize our notion of a scope-taker to make sense
of the idea that pronouns also \textit{scope}.

In \citeauthor{barkerShan2015}'s system, pronouns take scope in the following
way: they expect \textit{a proposition}, and they return an \textit{open
  proposition}.\sidenote{We can helpfully think of an open proposition in this
  framework as a proposition with anaphoric effects (i.e., environment sensitivity).}

In order for a \ac{qp} to bind a pronoun, it must first be
\textit{bind-shifted}. A bind-shifted \ac{qp} expects an open proposition, and
returns a proposition. Successful binding is illustrated in figure \ref{fig:cont1}.

\begin{figure}
\caption{Successful binding}\label{fig:cont1}
\begin{forest}
  [{$\type{t}$}
  [{$\type{\semtower{t}{e}}$\\$\ml{S}$},edge label={node[midway,left,font=\scriptsize]{$↓$}}
    [{$\type{\tower{t}{e → t}{e}}$} [{$\type{\semtower{t}{e}}$\\every boy},edge label={node[midway,left,font=\scriptsize]{$B$}}]]
    [{$\type{\tower{e → t}{t}{e → t}}$\\$\ml{S}$}
      [{$\type{\semtower{e → t}{e → e → t}}$\\loves}]
      [{$\type{\tower{e → t}{t}{e}}$} [{his mother},roof]]
    ]
  ]]
\end{forest}
\end{figure}

Putting mechanisms for inverse scope to one side, \ac{wco} follows
straightforwardly from this system. Since both pronouns and bind-shifted
\acp{qp} are scope-takers, for the pronoun to be bound, the \ac{qp} has to be
evaluated first. Scope can feed binding, but the \ac{qp} must precede the
pronoun, since \textit{evaluation order mirrors linear order}.\sidenote{One of
  the virtues of continuation semantics is that it straightforwardly accounts
  for scope out of \ac{dp} without requiring \textit{movement} out of \ac{dp}.}

Continuation semantics includes mechanisms for subverting the linear bias
(namely, higher-order continuations), in
order to account for inverse scope.

With mechanisms for inverse scope in the picture, things become a little less neat. \citet{barkerShan2015} must
stipulate that \textit{lower} -- the operation via which continuized meaning are
collapsed into ordinary meanings -- is rigidly typed. If we assume that
\textit{internal lower} is derived as lifted \textit{lower}, this also has
consequences for the type of internal lower:

\begin{multicols}{2}
$$↓ : \type{\tower{a}{t}{t} → a}$$
\columnbreak
$$⇊ : \type{\semtower{b}{\tower{a}{t}{t}} → \semtower{b}{a}}$$
\end{multicols}

This move basically guarantees, via a syntactic stipulation, that in order for a
bind-shifted \ac{qp} to bind a pronoun, it must take scope at the same
tower-story as the pronoun. If it takes scope on a high level, then the
resulting meaning cannot ultimately be lowered by a rigidly typed
\textit{lower}.

An unsuccessful attempt at getting internal lift to feed binding is illustrated
in figure \ref{fig:cont2}.

\begin{marginfigure}
\caption{Unsuccessful binding (\ac{wco})}\label{fig:cont2}
\begin{footnotesize}
\begin{forest}
  [{\xmark}
  [{$\type{\tower{t}{e → t}{e → t}}$},fill=yellow
  [{$\type{\tower{t}{e → t}{\tower{e → t}{t}{t}}}$\\$\ml{S}$},edge label={node[midway,left,font=\scriptsize]{$⇊$}}
    [{$\type{\semtower{t}{\tower{e → t}{t}{e}}}$},draw=red [{$\type{\tower{e → t}{t}{e}}$\\his mother},edge label={node[midway,left,font=\scriptsize]{$↑$}}]]
    [{$\type{\tower{t}{e → t}{\semtower{t}{e → t}}}$\\$\ml{S}$}
      [{$\type{\semtower{t}{\semtower{t}{e → e → t}}}$\\loves$^{↑ ∘ ↑}$}]
      [{$\type{\tower{t}{e → t}{\semtower{t}{e}}}$},draw=red [{every boy},edge label={node[midway,left,font=\scriptsize]{$⇈ ∘ B$}}]]
    ]
  ]
  ]]
\end{forest}
\end{footnotesize}
\end{marginfigure}

What's crucial here is that both lower and internal lower are rigidly typed.

Just how satisfying is this as an explanation though? If we look at what lower
actually \textit{does}, there's no intrinsic reason why it should be so rigidly
typed.

As shown in (\ref{def:lower}), all that lower does is feed its sole argument the
identity function. On a maximally polymorphic definition, therefore, the
argument need only be of type $\type{(a → a) → b}$.


\ex Lower (maximally polymorphic ver.)\\
$m^{↓} ≔ m id$\label{def:lower}
\xe

A maximally polymorphic lower could save the \ac{wco}-violating derivation in
figure \ref{fig:cont2}.

Based on what lower \textit{does}, there's no strong \textit{semantic}
motivation for making it rigidly typed. Therefore, despite the initial
conceptual appeal of \citeauthor{barkerShan2015}'s system, its success
ultimately rests on what looks like a syntactic stipulation.

Maybe we can do better. One recent attempt to derive \ac{wco} from independently
motivated semantic mechanisms is \citeauthor{chierchia2020}'s \textit{dynamic}
account of \ac{wco}.

\citeauthor{chierchia2020}'s account bears a family resemblance to that of
\citeauthor{barkerShan2015}. It too attempts to derive \ac{wco} from a semantic theory
with a built-in left-to-right bias.

\section{Chierchia: \textsc{wco} as a reflex of the dynamics of anaphora}

\citet{chierchia2020} develops a theory of \ac{wco} based on \textit{dynamic
  semantics}.

Much like continuation semantics, \ac{ds} is a semantic theory with a
\enquote{built-in} left-to-right bias.

\subsection{Dynamic semantics}

\ac{ds} is one of the most empirically successful theories of anaphora
(\citealt{heim1982,groenendijk_dynamic_1991,dekker1994}, a.o.) and
presupposition projection (\citealt{heim1983,beaver_presupposition_2001}, a.o.).
It has also been extended to a variety of other phenomena, including epistemic
modality, exhaustification, and more.

Crowning achievements of \ac{ds} include analyses of the following
phenomena:\sidenote{The (b) examples are included to briefly show that the
  phenomena under consideration exhibit a \textit{left-to-right asymmetry}, thus
motivating a dynamic treatment.}

\begin{itemize}

  \item Presupposition projection.\sidenote{Approaches to dynamic semantics are
    split as to whether they collapse presupposition satisfaction and anaphora
    resolution (see, e.g., \citealt{vanDerSandt1992}) or not
    (\citealt{heim1983}).

    The dynamic semantics ultimately adopted by \citeauthor{chierchia2020}
    follows the latter tradition. This won't be so important for the purposes of
    this class, but will be relevant when we talk about presupposition, starting
    from next week!
    }

    \pex
    \a{}[Ka visited Rome last summer]$^{α}$,\\
    and [she visited Rome again]$_{α}$
    this summer.
    \a\ljudge{\#}{}[Ka visited Rome again]$_{α}$,\\
    and [she visited Rome last summer]$^{α}$.
    \xe

    \item Donkey anaphora.

    \pex
    \a Every farmer who owns a donkey$^{3}$ treasures it$_{3}$.
    \a\ljudge{?}Every farmer who owns it$_{3}$ treasures a donkey$^{3}$.
    \xe

    \item Cross-sentential anaphora.

    \pex\label{ex:dy}
    \a A man$^{1}$ walked in. He$_{1}$ sat down.\label{ex:dy1}
    \a\ljudge{*}He$_{1}$ sat down. A man$_{1}$ walked in.\label{ex:dy2}
    \xe

 \end{itemize}

    Dwelling on cross-sentential anaphora, the contrast in (\ref{ex:dy}) is
    clearly reminiscent of a \ac{wco} effect.

    As we'll see, orthodox dynamic
    semantics doesn't by itself explain \ac{wco}, once quantificational scope is in the picture (see
    \citealt{charlow2019static} for discussion of this point), but
    \citeauthor{chierchia2020}'s basic intuition is to build a theory of
    \ac{wco} based on this contrast.

    In the next section, we'll introduce dynamic semantics by constructing an
    orthodox fragment that can account for cross-sentential anaphora. We'll move
    on to show how it fails to capture \ac{wco}, before moving on to
    \citeauthor{chierchia2020}'s modification.

\end{itemize}

\subsection{A Heimian fragment}

Sentential meanings in \ac{ds} (\citealt{heim1982}, \citealt{groenendijk_dynamic_1991},
\citealt{chierchia_dynamics_1995}), have two essential components:

\begin{itemize}

    \item An input-output asymmetry -- sentences denote \textit{instructions} to
    change the input context (see especially \citealt{heim1982}).

    \item Indeterminacy -- certain expressions may induce an
    \textit{indeterminate} output; an input can potentially be mapped to multiple outputs.

\end{itemize}



To model this formally, many theories of \ac{ds} model sentence meanings as \textit{relations between assignments} (equivalently: functions
from assignments, to sets of assignments).\sidenote[][-10\baselineskip]{In actual fact, worlds need
  to enter the picture too -- \citet{heim1982} models sentence meanings as
  relations between \textit{assignment-world pairs}. This is independently
  necessary in order to account for the presupposition projection facts. Since
  we'll only be looking at the dynamic account of anaphora resolution, it will
  be harmless to ignore this aspect of the theory.}

\acp{dp} introduce \acp{dr}, modeled as variables; indefinites, unlike definites
induce \textit{indeterminacy}, concerning the identity of the \ac{dr}. This is
illustrated schematically in figure \ref{fig:schema}.\sidenote[][-15ex]{One way of
  thinking about this: definites induce a \textit{functional} relation between
  assignments -- every input assignment is mapped to a unique output assignment,
whereas indefinites induce a \textit{non-functional} relation between
assignments -- each input assignment can mapped to one or more output assignments.}

\begin{figure}
\caption{Relations between assignments}\label{fig:schema}
\begin{forest}
  [{$ω$} [{Roger$^{n}$ arrived late.} [{$ω^{[n ↦ \ml{roger}]}$}]]]
\end{forest}
%
\begin{forest}
  [{$ω$} [{A linguist$^{n}$ arrived late}
    [{$ω^{[n ↦ \ml{kai}]}$}]
    [{$ω^{[n ↦ \ml{roger}]}$}]
    [{$ω^{[n ↦ \ml{sabine}]}$}]
    [{$ω^{[n ↦ \ml{athulya}]}$}]
    [{$ω^{[n ↦ \ml{martin}]}$}]
  ]]
\end{forest}
\end{figure}

\textit{Assignments} are functions from variables to individuals; as is
standard, we'll represent the set of variables as $ℕ$:\sidenote{We'll use
  $\type{o}$ as the type of assignments to distinguish between assignments used
  in a static setting.}

\ex Type of assignments\\
$\type{o} ≔ \type{n → e}$
\xe

\citeauthor{chierchia2020} assumes that assignments are \textit{partial}
functions.\sidenote{See also \citet{rothschildMandelkern2017esslli} for a
  \ac{ds} using partial assignments.} That is to say, an assignment
may only be defined for certain indices. The following are all valid assignments:

  \begin{multicols}{3}
  $$\left[
      1 ↦ \ml{roger}
    \right]$$
  \columnbreak
  $$\left[\begin{aligned}[c]
      &1 ↦ \ml{roger}\\
      &3 ↦ \ml{martin}
    \end{aligned}\right]$$
  \columnbreak
  $$\left[\begin{aligned}[c]
      &4 ↦ \ml{kai}\\
      &5 ↦ \ml{athulya}\\
      &7 ↦ \ml{sabine}\\
    \end{aligned}\right]$$
\end{multicols}

In order to characterize a dynamic sentential meaning, we define a type
constructor $\type{T}$ to abbreviate relations between assignments:

\ex Type of \acp{ccp}\\
$\type{T} ≔ \type{o → o → t}$
\xe

Here are some example sentence meanings:

Definites induce deterministic updates.

\ex
$\eval{Roger$^1$ arrived late} = λ ωω' . ω \stackrel{1/\ml{roger}}{=} ω' ∧ \ml{arrived-late roger}$\sidenote[][-15\baselineskip]{$ω \stackrel{n/x}{=} ω'$ is defined iff $ω_{n}$ is
  \textit{un}defined, and is true just in case $ω'$ differs from $ω$ at most in
  what $n$ is mapped to.


\citeauthor{heim1982}'s \textit{novelty condition} is essentially built into the
rule for \ac{dr} introduction.}\hfill$\type{T}$
\xe

  \begin{marginfigure}
    \centering
    \caption{Deterministic update}
    \begin{forest}
  [{$ω$} [{Roger$^{1}$ arrived late.} [{$ω^{[1 ↦ \ml{roger}]}$}]]]
\end{forest}
  \end{marginfigure}

Indefinites induce non-deterministic updates:

\ex $\eval{A linguist$^{1}$ arrived late} = λ ωω' . ∃x[ω \stackrel{1/x}{=} ω' ∧ \ml{arrived-late} x]$\hfill$\type{T}$
\xe

\begin{marginfigure}
\centering
\caption{Non-deterministic update}
\begin{forest}
  [{$ω$} [{A linguist$^{1}$ arrived late}
    [{$ω^{[1 ↦ \ml{kai}]}$}]
    [{$ω^{[1 ↦ \ml{roger}]}$}]
    [{$ω^{[1 ↦ \ml{sabine}]}$}]
  ]]
\end{forest}
\end{marginfigure}

We can get back an \enquote{ordinary} sentential meaning from a \ac{ccp} by
existentially closing the output assignment, as defined in (\ref{def:dyclo}).


\ex Dynamic closure (def.)\\
$m^{↓}} ≔ λ ω . ∃ω'[m ω ω']$\label{def:dyclo}\hfill$\dyclo : \type{T → o → t}$
\xe

If we apply dynamic closure to a non-deterministic update, we get classical,
existential truth-conditions; it will return true for any input assignment $ω$
just in case \textit{a linguist arrived late}.

How do we build up \acp{ccp} compositionally? \citeauthor{chierchia2020} assumes
that predicates are fundamentally Montagovian (i.e., functions of type
$\type{e → t}$):

\ex
$\eval{swim} ≔ λ x . \ml{swim} x$\hfill$\type{e → t}$
\xe



Predicates are lifted into a dynamic setting by a type-shifter
\textit{dynamic lift}; d-lift takes a function from an individual to a
truth-value, and shifts it into a function from an individual to a \ac{ccp} --
specifically, a dynamic \textit{test}.\sidenote[][-7\baselineskip]{
A different way of generalizing this to $n-$place predicates is by giving $Δ$ the following definition:

\ex
$m^{Δ} ≔ λ k . λ ωω' . ω = ω' ∧ m k$\\
$Δ: \type{((a → t) → t) → (a → t) → t}$
\xe

\Citeauthor{chierchia2020}'s d-lift can be derived as follows, using the (by now
very familiar) continuation semantics operations:

\ex
$λ x . (x^{↑} \ml{S} f^{Δ ∘ ↑})^{↓}$
\xe
}

\ex
Dynamic lift (def.)\\
$f^{\dlift} ≔ λ x . λ ω . λ ω' . ω = ω' ∧ f x$\hfill$\dlift : \type{(e → t) → e
  → T}$
\xe

\begin{tcolorbox}
\textbf{Exercise}
\tcblower
\citeauthor{chierchia2020} defines dynamic lift in such a way that it only can
apply to one-place predicates. This is not insignificant -- see the discussion
of event semantics later on. It is however trivial to generalize to d-lift to
$n$-place predicates.

Generalize \textit{dynamic lift} to $n$-place predicates by giving a recursive
definition a la \cite{parteeRooth}.
\end{tcolorbox}

Tests don't do anything interesting to input contexts. If we d-lift
\textit{swims} and apply it to \textit{John} we'll get back a test that relates
the input context $ω$ to itself, iff \textit{John swims} is true.

In orthodox dynamic
fragments, all of the interesting dynamic action is triggered by arguments --
specifically, pronouns and indefinites.


A pronoun indexed $n$ expects a dynamic predicate $k$ as its input, and returns
a \ac{ccp} -- a function from an input assignment $ω$ to the result of feeding
$ω_{n}$ into $k$, \textit{re-}saturated with $ω$.\sidenote[][-15\baselineskip]{The type signature of
a pronoun betrays the fact that, in this dynamic grammar, pronouns are
\textit{scope-takers}, and in fact, we can abbreviate a pronominal meaning using
tower notation:

\ex
Pronouns (tower def.)\\
$\ml{pro}_{n} ≔ \semtower{λ ω . ([] ω)}{ω_{n}}$
\xe

Interestingly, this is what we get if we apply the \textit{bind} of the
\texttt{Reader} monad to the static entry for a pronoun.

\ex Pronoun (static def.)\\
$\ml{pro}_{n} ≔ λω . ω_{n}$\\
\phantom{,}\hfill$\type{o → e}$
\xe

\ex~ Bind of \texttt{Reader} (def.)\\
$m^{⋆} ≔ λ k . λ ω . k (m ω) ω$\\
\phantom{,}\hfill$\type{(o → a) → (a → o → b) → o → b}$
\xe
}

\ex
Pronouns (def.)\\
$\ml{pro}_{n} ≔ λ k . λ ω . k ω_{n} ω$\sidenote{
\citeauthor{chierchia2020} actually posits a syncategorematic rule for composing
pronouns and dynamic predicates -- instead, I've built what
\citeauthor{chierchia2020}'s rule does into the meaning of the pronoun.
}\hfill$\ml{pro}_{n} : \type{(e → T) → T}$
\xe

Pronouns now may compose with d-lifted predicates via \ac{fa}, as illustrated
in figure \ref{fig:pro}:

\begin{figure}
\caption{Pronouns in a dynamic fragment\\\enquote{He$_{3}$ swims}}\label{fig:pro}
\begin{forest}
  [{$λωω' . ω = ω' ∧ \ml{swim} ω_{3}$},fill=yellow
    [{$λ k . λ ω . k ω_{3} ω$\\he$_{n}$}]
    [{$λ x . λ ωω' . ω = ω' ∧ \ml{swim} x$} [{swim},edge label={node[midway,left,font=\scriptsize]{$\dlift$}}]]
  ]
\end{forest}
\end{figure}

The result is a dynamic \textit{test}, that saturates the argument of
\textit{swim} with whatever the input assignment $ω$ maps to pronominal index
$3$ to.

In an orthodox dynamic fragment (\citealt{heim1982,groenendijk_dynamic_1991}), indefinites introduce
\acp{dr}.\sidenote{\citeauthor{chierchia2020} will ultimately reject this
  assumption, but it will be useful to consider his claims in light of the
  standard theory.}

\ex
Indefinites (Heimian def.)\\
$\ml{someone}_{n} ≔ λ k . λ ωω' . ∃x,w''[ω \stackrel{n/x}{=} ω'' ∧ k x ω'' ω']$\\
\phantom{,}\hfill$\ml{someone}_{n} \type{(e → T) → T}$
\xe

In figure \ref{fig:indef}, we show how a Heimian indefinite composes in a
dynamic fragment. The result maps each input assignment $ω$ to (the
characteristic function of) a \textit{set} of assignments $ω'$, s.t., $ω'_{n}$
is a swimmer.

\begin{figure}
\caption{Heimian indefinites in a dynamic fragment\\\enquote{Someone$_{7}$ swims}}\label{fig:indef}
\begin{forest}
  [{$λ ωω' . ∃x[ω \stackrel{7/x}{=} ω' ∧ \ml{swim} x]$},fill=yellow
  [{$λωω' . ∃x,w''[ω \stackrel{7/x}{=} ω'' ∧ ω'' = ω' ∧ \ml{swim} x]$},edge label={node[midway,left,font=\scriptsize]{equiv}}
    [{$λ k . λ ωω' . ∃x,w''[ω \stackrel{7/x}{=} ω'' ∧ k x ω'' ω']$\\someone$_{7}$}]
    [{$λ x . λ ωω' . ω = ω' ∧ \ml{swim} x$} [{swim},edge label={node[midway,left,font=\scriptsize]{$\dlift$}}]]
  ]]
\end{forest}
\end{figure}

A famous design feature of \ac{ds} is an account of cross-sentential
binding, as in the following famous examples:

\pex
\a Someone$^{1}$ walked in and he$_{1}$ sat down.\label{ex:seq1}
\a Someone$^{1}$ walked in. he$_{1}$ sat down.\label{ex:seq2}
\xe

In \ac{ds}, conjunction -- as in (\ref{ex:seq1}) -- is treated as a
special case of \textit{discourse sequencing} (\ref{ex:seq2}).

Discourse sequencing is an operation on \acp{ccp}:

\ex Dynamic sequencing (def.)\\
$m ; n ≔ λω . λ ω' . ∃ω''[m ω ω'' ∧ n ω'' ω']$\hfill$(;): \type{T → T → T}$\label{def:conj}
\xe

An illustration of how cross-sentential anaphora works in a Heimian fragment is
given in figure \ref{fig:anaph}: sequencing the \acp{ccp} gives rise to a
\ac{ccp} that relates $ω$ and $ω'$ just in case $ω_{4}$ is undefined and and
$ω'_{4}$ walked in and sat down.

\begin{figure}
\caption{Cross-sentential anaphora in a Heimian fragment}\label{fig:anaph}
\begin{forest}
  [{$λ ωω' . ∃ω''[(∃x[ω \stackrel{4/x}{=} ω'' ∧ \ml{walked-in} x]) ∧ (ω'' = ω' ∧ \ml{sat-down} ω''_{4})]$},fill=yellow,name=landing
  site
    [{$λ ωω' . ∃x[ω \stackrel{4/x}{=} ω' ∧ \ml{walked-in} x]$},draw=red [{Someone$^{4}$ walked in},roof]]
    [{...}
      [{$λ n . λ m . λωω' . ∃ω''[m ω ω'' ∧ n ω'' ω']$\\$;$}]
      [{$λωω' . ω = ω' ∧ \ml{sat-down} ω_{4}$},name=trace,draw=red [{he$_{4}$ sat down},roof]]
    ]
  ]
\end{forest}
\end{figure}

Why is \ac{ds} promising as a starting point for a theory of \ac{wco}? Recall the contrast below, reminiscent of \ac{wco}:

\pex
\a Someone$^{4}$ walked in and he sat down.
\a\ljudge{*}He$_{4}$ walked in and someone$^{4}$ sat down.\label{ex:cat2}
\xe

Just so long as \textit{someone} takes scope within its containing sentence,
\ac{ds} captures this contrast, by virtue of the left-to-right bias
built into the definition of discourse sequencing.

If we try to compute the \ac{ccp} for (\ref{ex:cat2}), the result is guaranteed
to be undefined. This is because, if the input assignment $ω$ is defined for
$4$, it can't also be \textit{undefined} for $4$, as is required by the meaning
contributed by the indefinite.

\ex
$λ ωω' . ∃ω''[ω = ω'' ∧ \ml{sat-down} ω_{4} ∧ (∃x[ω'' \stackrel{4/x}{=} ω' ∧ \ml{walked-in} x])]$
\xe

\ac{ds} doesn't by itself however capture \ac{wco} -- this is because,
independently, we need a mechanism that allows indefinites to \textit{take
  scope}; indefinites introduce discourse referents at their
scope site. We can therefore compute a bound reading for the following example,
by scoping the indefinite over the pronoun:

\begin{figure}
  \caption{Violating \ac{wco} in a Heimian fragment\\\enquote{He$_{4}$ wants to
      meet someone$^{4}$}}
  \begin{forest}
  [{$λ ωω' . ∃x[ω \stackrel{4/x}{=} ω' ∧ ω'_{4} \ml{want} (ω_{4} \ml{meet} x)]$},fill=yellow
    [{someone$_{4}$},name=intermediate]
    [{$λ x . λ ωω' . ω = ω' ∧ ω_{4} \ml{want} (ω_{4} \ml{meet} x)$}
    [{$λ x$}]
    [{$λ ωω' . ω = ω' ∧ ω_{4} \ml{want} (ω_{4} \ml{meet} x)$}
      [{he$_{4}$}]
      [{$λ y . λωω' . ω = ω' ∧ y \ml{want} (y \ml{meet} x)$} [{$λ y . y$ wants to meet $x$},edge label={node[midway,left,font=\scriptsize]{$Δ$}},name=trace]]
    ]
  ]]
  {
    \draw[semithick, dashed, ->] ([xshift=4em]trace.south) to[out=south,in=south] (intermediate); % e
  }
  \end{forest}
\end{figure}

Intuitively, a problematic feature of \ac{ds} in this regard is that it ties
together \ac{dr} introduction with quantificational scope.


\subsection{The Dynamic Prediction Principle}

At the heart of \citeauthor{chierchia2020} account of \ac{wco} is an apparently
minor modification to orthodox dynamics, with far reaching consequences: the
\acf{dpp}, stated in (\ref{def:dpp}).

\ex
The \acf{dpp}\\
\acp{dr} can only be introduced by predicates.\hfill (\citealt[p.\,32]{chierchia2020})\label{def:dpp}
\xe

\citeauthor{chierchia2020}'s innovation is to posit a second way of lifting
predicates into a dynamic setting: \textit{\ac{dr}-lifting}.\sidenote{If
  you try to generalize \ac{dr}-lift to $n$-place predicates, you'll
  find that it can't be done in quite the same way as for d-lift. As an
  exercise, try to figure out why this.}

\ex
\ac{dr}-lift (def.)\\
$f^{Δ_{n}} ≔ λ x . λ ω . λ ω' ω \stackrel{n/x}{=} ω' ∧ f x$\hfill$Δ_{n}: \type{(e → t) → e
  → T}$
\xe

Introducing \acp{dr} then, is no longer the job of \textit{indefinites}, but
rather the job of a \ac{dr}-lifted predicate.

What do indefinites do then? For \citeauthor{chierchia2020}, they're just
type-lifted first-order quantifiers.\sidenote{Looking at the definition in
  (\ref{def:someone}), you may be wondering how \textit{someone} binds its
  trace. \citeauthor{chierchia2020} does something rather sneaky here, which
  will be important later. For now, assume that it just works.}

\ex
Dynamic existential quantification (def.)\\
$\ml{someone}_{n} m ≔ λωω' . ∃x_{n}[m ω ω']$\hfill$\type{\ml{someone}_{n}: \type{T → T}}$\label{def:someone}
\xe

\textit{Someone} saturates the argument that a \ac{dr} was introduced relative
to, and cross-sentential anaphora proceeds as usual.

\begin{figure}
  \centering
  \caption{Example derivation}
  \begin{forest}
    [{$λ ωω' . ∃ω''[∃x[ω \stackrel{1/x}{=} ω'' ∧ \ml{walked-in} x] ∧ ω'' = ω' ∧ \ml{sat-down} ω''_1]$},fill=yellow
    [{$λ ωω' . ∃x[ω \stackrel{1/x}{=} ω' ∧ \ml{walked-in} x]$},draw=red
      [{someone$_{2}$}]
      [{$λ ωω' . ω \stackrel{1/x}{=} ω' ∧ \ml{walked-in} x_{2}$}
      [{$t_{2}$}]
      [{$λ x . λ ωω' . ω \stackrel{1/x}{=} ω' ∧ \ml{walked-in} x$} [{walked in},edge label={node[midway,left,font=\scriptsize]{$Δ_{1}$}}]]]]
    [{...}
    [{$;$}]
      [{$λ ωω' . ω = ω' ∧ \ml{sat-down} ω_1$},draw=red
        [{$λ k . λ ω . k ω_1 ω$\\he$_1$}]
        [{$λ y . λ ωω' . ω = ω' ∧ \ml{sat-down} y$} [{sat-down},edge label={node[midway,left,font=\scriptsize]{$Δ$}}]]
      ]
    ]]
  \end{forest}
\end{figure}

So far, we've constructed a system which replicates the basic results of
orthodox dynamic semantics, but with a different compositional regime.

\subsection{Accessibility}


In the previous section, we only gave definitions for \textit{dynamic
  conjunction/discourse sequencing} and the static first order existential.

\citeauthor{chierchia2020} adopts the standard dynamic definitions for the other
logical operators.

Negation is taken to be \textit{externally
  static}; any \acp{dr} introduced in the scope of negation are
subsequently wiped out.

\ex
Dynamic negation (def.)\\
$¬ m ≔ λ ω . λω' . ω = ω' ∧ ¬ (m^{↓} ω)$\hfill$¬ : \type{T → T}$
\xe

Externally static negation predicts the impossibility of binding in the
following:\sidenote{
  You might wonder about what we predict for a sentence such as the following,
  where the indefinite takes wide scope over negation:

  \ex
  It's not true that someone walked in.\\He sat down.
  \xe

  This will be important later.
}

\ex
\ljudge{*}It's not the case that anyone$^{1}$ walked in. He$_{1}$ sat down.
\xe

To see why, first consider the prejacent of negation, with \ac{dr}-lift applied
to the predicate:\sidenote{We simplify here and assume that \ac{npi}
  \textit{any} is just an existential licensed in the scope of negation.}

\ex
$\eval{anyone walked in} = λ ωω' . ∃x[ω \stackrel{1/x}{=} ω' ∧ \ml{walked-in} x]$\label{first-conj}
\xe

Applying dynamic negation to the above \ac{ccp} \textit{existentially closes}
the output assignment, thereby rendering it dynamically inert. The resulting
\ac{ccp} is a dynamic test, and asserts that there is no way of extending the
input assignment s.t. $1$ is mapped to someone who walked in (in other words,
nobody walked in).

\ex
$¬ (\ref{first-conj}) = λ ωω' . ω = ω' ∧ ¬ ∃ω'',x[ω \stackrel{1/x}{=} ω'' ∧ \ml{walked-in} x]$
\xe

Sequencing this \ac{ccp} with the second conjunct will clearly not give rise to anaphora.

The remainder of the logical operations can be defined via first-order
equivalent via dynamic conjunction, negation, and existential quantification.
All are defined as operations on \acp{ccp}.

\ex Dynamic implication (def.)\\
$m → n ≔ ¬ (m ; ¬ n)$\hfill$(→) : \type{T → T → T}$
\xe

\ex~ Dynamic disjunction (def.)\\
$m ∨ n ≔ ¬ (¬ m ; ¬ n)$\hfill$(∨): \type{T → T → T}$
\xe

\ex~ Dynamic universal quantifier (def.)\\
$\ml{everyone}_{n} m ≔ ¬ ∃x_{n} (¬ m)$\hfill$\type{T → T}$
\xe

Famously, this way of dynamicizing the logical connectives gives rise to the
following accessibility hierarchy in complex sentences:

\ex Accessibility (def.)\\
A is \textit{accessible} to B if a \ac{dr} active in A can covary with a pronoun
in B.
\xe

    Accessibility in conjunctive sentences: [A and B]\sub{S}

    \begin{itemize}

        \item A is accessible to B (but not vice versa).

        \item B is accessible to whatever is conjoined with S.

    \end{itemize}

        \ex
        A man$^{1}$ walked in, and he$_{1}$ sat down. He$_{1}$ stood up again
        soon after.
        \xe


    Accessibility in conditional sentences: [if A then B]\sub{S}

    \begin{itemize}

        \item A is accessible to B (but not vice versa).

        \item A, B are \textit{not} accessible to what is conjoined with S.\sidenote{In dynamic semantics, conditional sentences are
        internally dynamic, but externally static.}

    \end{itemize}

            \ex
        \ljudge{\#}If someone$^{1}$ won the lottery, they$_{1}$ became rich. I
        shook their$_{1}$ hand.
        \xe


    Accessibility in negative sentences: [not A]\sub{S}

    \begin{itemize}

        \item Nothing in A is accessible to what is conjoined with S.

    \end{itemize}

            \ex
        \ljudge{*}It's not the case that anyone$^{1}$ sat down. He$_{1}$ walked in.
        \xe

    Accessibility in disjunctive sentences: [A or B]\sub{S}

    \begin{itemize}

        \item A is not accessible to B, nor is B to A.

        \item Neither A not B is accessible to what is conjoined with S.\sidenote{Dynamic disjunction is both internally static and
        externally static.}

    \end{itemize}

            \ex
        \ljudge{\#}Either Mary has a new dog$^{1}$, or I petted it$_{1}$.
        \xe


\subsection{Enter events}

So far, we've constructed a fragment that only accommodates one-place
predicates. This is actually by design -- \citeauthor{chierchia2020} argues that
such a system has a natural bed-fellow in neo-Davisonian event
semantics.\sidenote{
We'll take the type of an event to be $\type{v}$.
}

Traditions in event semantics:\sidenote{Not discussed here is the hybrid
  approach due to \cite{kratzer_severing_1996}, where only the external argument
is introduced by a distinct thematic head.}

Davidsonian:

\ex
$\eval{love} ≔ λ exy . \ml{exp} e = y ∧ \ml{th} e = x ∧ \ml{love} e$\hfill$\type{v → e → e → t}$
\xe

Neo-Davidsonian (\citealt{castaneda1967,parsons_events_1990}):

\ex
$\eval{love} ≔ λ e . \ml{love} e$\hfill$\type{v → t}$
\xe

According to the neo-Davidsonian approach, all arguments are severed, and
instead introduced by thematic role heads (the compositional regime adopted here
is after \citealt{champollion_interaction_2015}):\sidenote{See, e.g. \citealt{ahn2016} and
\citealt{elliottDiss} for independent evidence for this position from different domains.}

Thematic role heads take individuals and return properties of events --
$\ml{THEME}$ returns a function from $x$ to events of which $x$ is the \textit{theme}.

\ex
$\ml{THEME} ≔ λ x . λ e . \ml{th} e = x$\hfill$\type{e → v → t}$
\xe

Abstracting away from dynamics for a moment, the composition of a simple
sentence in a neo-Davidsonian setting can proceed via \ac{pm}:

\begin{figure}
\centering
\caption{composition in a neo-Davidsonian event semantics}
\begin{forest}
    [{$λ e . \ml{exp} e = \ml{john} ∧ \ml{th} e = ιx[\ml{cat} x] ∧ \ml{love} e$\\\ac{pm}},fill=yellow
    [{$λ e . \ml{exp} e = \ml{john}$}
      [{$\ml{EXP}$}]
      [{DP\\John}]
    ]
        [{$λ xe . \ml{th} e = ιx[\ml{cat} x] ∧ \ml{love} e$\\\ac{pm}}
        [{$λ e . \ml{th} e = ιx[\ml{cat} x]$}
          [{$\ml{THEME}$}]
          [{DP} [{the cat},roof]]
        ]
          [{$λ e . \ml{love} e$\\love}]
      ]
    ]
\end{forest}
\end{figure}

Note that, since verbs denote one place predicates, they can be \ac{dr}-lifted.

Since the event argument of a verb is (by stipulation) existentially closed,
this accounts for the possibility of eventive \acp{dr}, as in the following
example:

\ex
It rained$^{4}$. It$_{4}$ was heavy.
\xe

We can assume the following \ac{lf}:

\begin{figure}
  \centering
  \caption{Eventive \acp{dr}}
  \begin{forest}
    [{$λ ωω' . ∃ω''[(∃e[ω \stackrel{4/e}{=} ω'' ∧ \ml{rain} e]) ∧ ω'' = ω' ∧ \ml{heavy} ω_{4}]$},fill=yellow
    [{$λωω . ∃e[ω \stackrel{4/e}{=} ω' ∧ \ml{rain} e]$},draw=red
      [{$∃$}]
      [{$λ eωω' . ω \stackrel{4/e}{=} ω' ∧ \ml{rain} e$} [{rain},edge label={node[midway,left,font=\scriptsize]{$Δ_{4}$}}]]
    ]
      [{...}
        [{$;$}]
        [{$λ ωω' . ω = ω' ∧ \ml{heavy} ω_{4}$},draw=red [{it was heavy},roof]]
      ]
    ]
  \end{forest}
\end{figure}

The discussion in the paper is quite confusing at this point, but
\citeauthor{chierchia2020} seems to assume that \ac{dr}-lift extends
straightforwardly to thematic argument-introducing heads. This is in fact not
the case -- we have to generalize \ac{dr}-lift to predicates of type
$\type{e → v → t}$.\sidenote{As far as I can see, the compositional details of
  the system as laid out by \citeauthor{chierchia2020} at this point are
  incoherent, but easily fixed.}

\ex Thematic \ac{dr}-lift (def.)\\
$f^{Δ_{n}} ≔ λ x . λ e . λωω' . ω \stackrel{n/x}{=} ω' ∧ f x e$\hfill$Δ_{n}:\type{(e → v → t) → e → v → T}$
\xe

We furthermore must assume that \textit{dynamic} \ac{pm} is a freely available
semantic composition rule -- dynamic \ac{pm} is just like ordinary \ac{pm}, only
instead of conjoining the inner propositional value, we do dynamic sequencing.

\ex Dynamic \ac{pm} (def.): $\eval{\begin{aligned}[c]
    \begin{forest}
    [
      [{$m_\type{a → T}$}]
      [{$n_\type{a → T}$}]
    ]
    \end{forest}
  \end{aligned}} ≔ λ x_{\type{a}} . m x ; n x$
\xe

If we return to our simple example, we can now apply \ac{dr}-lift every step of
the way, as in figure \ref{ex:inter}:

\begin{figure}
\centering
\caption{Dynamic neo-Davidsonian event semantics}\label{ex:inter}
\begin{forest}
  [{$λ e . \begin{aligned}[t]
      &(λωω' . ω \stackrel{3/j}{=} ω' ∧ \ml{exp} e = j)\\
      &;(λωω' . ω \stackrel{2/ιx[\ml{cat} x ω_{3}]}{=} ω' ∧ \ml{th} e = ιx[\ml{cat} x ω_{3}])\\
      &;(λωω' . ω \stackrel{1/e}{=} ω' ∧ \ml{love} e)\end{aligned}$\\d-\ac{pm}},fill=yellow
    [{$λ e . λωω' . ω \stackrel{3/j}{=} ω' ∧ \ml{exp} e = j$},draw=red
      [{...} [{$\ml{EXP}$},edge label={node[midway,left,font=\scriptsize]{$Δ_{3}$}}]]
      [{DP\\John}]
    ]
        [{d-\ac{pm}}
        [{$λ e . λωω' . ω \stackrel{2/ιx[\ml{cat} x ω_{3}]}{=} ω' ∧ \ml{th} e = ιx[\ml{cat} x ω_{3}]$},draw=red
          [{...} [{$\ml{THEME}$},edge label={node[midway,left,font=\scriptsize]{$Δ_{2}$}}]]
          [{DP} [{his$_{3}$ cat},roof]]
        ]
          [{$λ e . λωω' . ω \stackrel{1/e}{=} ω' ∧ \ml{love} e$},draw=red [{love},edge label={node[midway,left,font=\scriptsize]{$Δ_{1}$}}]]
      ]
    ]
\end{forest}
\end{figure}

What this essentially buys us is a sentence-internal accessibility hierarchy --
the subject and object are accessible to the verb, but not nice versa, and the
subject is accessible to the object, but not vice versa.

This can be leveraged in order to account for the basic cases of \ac{wco}.
Successful Q-binding is illustrated in figure \ref{fig:succ}; the subject is
accessible to the object, so the trace of the \ac{qp} comes to dynamically bind
the pronoun by \ac{dr}-shifting the thematic role head.

\begin{figure}
  \centering
  \caption{Successful Q-binding: \enquote{Everyone loves his cat}.}\label{fig:succ}
\begin{forest}
  [{$∀_{x} ∃_{e} \begin{aligned}[t]
      &(λωω' . ω \stackrel{3/x}{=} ω' ∧ \ml{exp} e = x)\\
      &;(λωω' . ω = ω' ∧ \ml{th} e = ιy[\ml{cat} y ω_{3}])\\
      &;(λωω' . ω = ω' ∧ \ml{love} e)\end{aligned}$},fill=yellow
  [{\ac{fa}}
  [{$∀x$\\everyone$_{x}$}]
  [{\ac{fa}}
  [{$∃_{e}$}]
  [{d-\ac{pm}}
    [{$λ e . λωω' . ω \stackrel{3/x}{=} ω' ∧ \ml{exp} e = x$},draw=red
      [{...} [{$\ml{EXP}$},edge label={node[midway,left,font=\scriptsize]{$Δ_{3}$}}]]
      [{DP\\$t_x$}]
    ]
        [{d-\ac{pm}}
        [{$λ e . λωω' . ω = ω' ∧ \ml{th} e = ιy[\ml{cat} y ω_{3}]$},draw=red
          [{...} [{$\ml{THEME}$},edge label={node[midway,left,font=\scriptsize]{$Δ$}}]]
          [{DP} [{his$_{3}$ cat},roof]]
        ]
          [{$λ e . λωω' . ω = ω' ∧ \ml{love} e$},draw=red [{love},edge label={node[midway,left,font=\scriptsize]{$Δ$}}]]
      ]
    ]]]]
\end{forest}
\end{figure}

Unsuccessful Q-binding is illustrated in figure \ref{fig:unsucc} -- since the object is \textit{not} accessible to the subject, discourse binding of the pronoun by the trace of the \ac{qp} fails.

\begin{figure}
  \centering
  \caption{Successful Q-binding: \enquote{Everyone loves his cat}.}\label{fig:unsucc}
\begin{forest}
  [{$∀_{x} ∃_{e} \begin{aligned}[t]
      &(λωω' . ω = ω' ∧ \ml{exp} e = ιy[\ml{cat} y ω_{3}])\\
      &;(λωω' . ω \stackrel{3/x}{=} ω' ∧ \ml{th} e = x)\\
      &;(λωω' . ω = ω' ∧ \ml{love} e)\end{aligned}$},fill=yellow
  [{\ac{fa}}
  [{$∀x$\\everyone$_{x}$}]
  [{\ac{fa}}
  [{$∃_{e}$}]
  [{d-\ac{pm}}
    [{$λ e . λωω' . ω = ω' ∧ \ml{exp} e = ιy[\ml{cat} y ω_{3}]$},draw=red
      [{...} [{$\ml{EXP}$},edge label={node[midway,left,font=\scriptsize]{$Δ$}}]]
      [{DP} [{his$_{3}$ cat},roof]]
    ]
        [{d-\ac{pm}}
        [{$λ e . λωω' . ω \stackrel{3/x}{=} ω' ∧ \ml{th} e = x$},draw=red
          [{...} [{$\ml{THEME}$},edge label={node[midway,left,font=\scriptsize]{$Δ_3$}}]]
          [{DP} [{$t_x$},roof]]
        ]
          [{$λ e . λωω' . ω = ω' ∧ \ml{love} e$},draw=red [{love},edge label={node[midway,left,font=\scriptsize]{$Δ$}}]]
      ]
    ]]]]
\end{forest}
\end{figure}

One thing we should be explicit about now -- just how \textit{do} quantifiers
bind their traces? In the paper, it's assumed that the meta-language is
something like first order logic -- traces of quantifiers denote \textit{first
  order} variables, and therefore as long as a trace is co-indexed with a
quantifier, it is bound -- traces of \acp{qp} are indirectly bound.

\citeauthor{chierchia2020} therefore assumes a completely different system of
indexation and binding specific to \acp{qp}. If we wanted to translate
\citeauthor{chierchia2020}'s approach into a theory in which binding is more
direct, we'd need to redefine the \ac{ccp} type constructor as follows:

\ex
$\type{T ≔ o → o → g → t}$
\xe

The need to distinguish between first order variables and genuine pronouns has
precedents in the dynamic literature\sidenote{\cite{dekker1994}}, but this is
arguably a conceptually unappealing aspect of the proposal.\sidenote{See
  \citeauthor{barkerShan2008}'s criticism of \ac{dmg} for relevant discussion.}

\subsection{Binding into adjuncts}

\citeauthor{chierchia2020} argues that, an advantage of incorporating events
into the system, is that his theory straightforwardly accounts for binding into
adjuncts\sidenote{At least, adverbial adjuncts that are interpreted as event
  modifiers -- see \cite{parsons_events_1990}.}

In event semantics, adverbials are simply interpreted as properties of events,
and are incorporated into the sentence via d-\ac{pm}, below existential closure.

\ex
$\eval{against John's will} = λ e . e \ml{against} \ml{will j}$\hfill$\type{v → t}$
\xe

Since, the adverbial is a monadic predicate, it can of course be d-lifted:

\ex
$\eval{against John's will}^{Δ} = λ e . λωω' . ω = ω' ∧ e \ml{against} \ml{will j}$\hfill$\type{v → T}$
\xe

Since adverbials are adjoined to the right typically, the system predicts that
\textit{objects are accessible to right-adjoined adverbials}. We can now account
for the following example straightforwardly:

\ex
John loves everyone against their will.
\xe

\begin{figure}
  \centering
  \caption{Binding into adjuncts}
  \begin{forest}
    [{$\type{T}$}
      [{$∀_{x}$\\everyone$_{x}$}]
      [{$\type{T}$}
        [{$∃_{e}$}]
        [{$\type{v → T}$}
        [{$\type{v → T}$}
          [{$\type{e → v → T}$} [{$\ml{EXP}$},edge label={node[midway,left,font=\scriptsize]{$Δ$}}]]
          [{DP\\John}]
        ]
        [{$\type{v → T}$}
        [{$\type{v → T}$}
        [{$\type{v → T}$}
          [{$\type{e → v → T}$},tikz={\node [draw=red,dashed,inner sep=0,fit
         to=tree]{};} [{$\ml{THEME}$},edge label={node[midway,left,font=\scriptsize]{$Δ_{3}$}}]]
          [{$t_{x}$}]
        ]
           [{$\type{v → T}$} [{loves},edge label={node[midway,left,font=\scriptsize]{$Δ$}}]]
        ]
          [{$\type{v → T}$},tikz={\node [draw=red,dashed,inner sep=0,fit
         to=tree]{};} [{against its$_{3}$ will},roof]]
        ]
        ]
      ]
    ]
  \end{forest}
\end{figure}

\section{Problems}

\subsection{A-movement}

A-movement bleeds \ac{wco}:

\ex
Every boy$_{x}^{1}$ seems to his$_{1}$ mother [$t_{x}$ to be happy].
\xe

Furthermore, A-movement out of an externally static environment feeds dynamic
binding:

\ex
Someone didn't leave. He sat down.
\xe

This leads to an unavoidable weakening of the \ac{dpp}:

\pex
The \acf{dpp} (refined ver.)\\
\acp{dr} can only be introduced by:
\a A-positions
\a predicates.
\xe

Theoretically, this means that the operation of \ac{dr}-lifting must be licensed
at derived predicates created by A-movement. This doesn't seem particularly
enlightening or explanatory.

\subsection{The problem of existentials}

Since \ac{dr}-introduction is performed by \textit{predicates} rather than arguments,
\citeauthor{chierchia2020} makes a bad prediction: Scoping an indefinite out of
an \textit{externally static} environment should fail to feed
binding.\sidenote{It's not necessary to use a specific indefinite to illustrate
  this point, but it does make the wide-scope reading more salient.}

This is clearly a bad prediction, as illustrated by (\ref{ex:indef})

\ex
It's not the case that [\fbox{a certain boy}$^{1}$ sat down]. He$_{1}$ left
hours ago.\label{ex:indef}
\xe

To see \textit{why} \citeauthor{chierchia2020} makes this prediction, it's
enough to consider the simple event-free fragment from the beginning.

First, consider the meaning of the prejacent of negation, with a \ac{dr}-shifted
predicate relative to $1$. This introduces a \ac{dr} corresponding to the trace
of the quantifier.

\ex
$\eval{$t_x$ sat down} = λωω' . \stackrel{1/x}{=} ω' ∧ \ml{sat-down} x$\label{step1}
\xe

Applying dynamic negation to the prejacent closes off the discourse referent:

\ex
$¬ (\ref{step1}) = λ ωω' . ω = ω' ∧ ¬ ∃ω''[ω \stackrel{1/x}{=} ω'' ∧ \ml{sat-down} x]$\label{step2}
\xe

Now, binding the trace with an existential quantifier fails to re-introduce the
\ac{dr} that was wiped out by dynamic negation.

\ex
$∃_{x} (\ref{step2}) = λ ωω' . ω = ω' ∧ ∃x[¬ ∃ω''[ω \stackrel{1/x}{=} ω'' ∧ \ml{sat-down} x]]$\label{step3}
\xe

\citeauthor{chierchia2020} concludes that (a) we should treat indefinites as
\textit{choice-functional variables}, and furthermore than (b) existential
closure of a choice-functional variable introduces a \ac{dr}. This is clearly a
weakening of the \ac{dpp}.

\pex
The \acf{dpp} (refined ver.)\\
\acp{dr} can only be introduced by:
\a A-positions
\a predicates.
\a existential closure of choice-functional variables.
\xe

The refined version of the theory predicts that \textit{indefinite scope feeds \ac{dr}-introduction}.

Note that this predicts that \textit{wide-scope} indefinites should obviate
\ac{wco}. \citeauthor{chierchia2020} argues that this is a good prediction based
on the following contrasts. As has been observed before, \textit{specific
  indefinites can obviate \ac{wco}}.

\pex
\a\ljudge{*}His$_{1}$ father hates a boy$^{1}$.
\a\ljudge{?}His$_{1}$ father hates $\set{\text{a boy I know}^{1} | \text{a certain
    boy}^{1} | \text{a friend of mine}^{1}}$
\xe

Unfortunately, this tying together of \ac{dr} introduction and \ac{wco}
obviation is going to lead to a fatal flaw in \citeauthor{chierchia2020}'s
theory. Anything that can introduce a \ac{dr} by scoping out of an externally
static environment must introduce a \ac{dr} at its \textit{scope site}; anything
that does so is predicted to obviate \ac{wco}.

\textit{wh}-expressions can introduce \acp{dr}:

\ex
Who$^{1}$ walked in? and, did they$_{1}$ sit down?
\xe

Furthermore, \textit{wh}-moving out of an externally static environment feed
\ac{dr} introduction:

\ex
Who$^{1}_{x}$ does nobody like $t_{x}$? and, where are they$_{1}$?
\xe

Even wh-in-situ can scope out of an externally static environment, feeding
\ac{dr} introduction:

\ex
Which boy$^{1}_{y}$ $t_{y}$ bought none of his friends which book$^{2}$?\\
and, why did he$_{1}$ hate it$_{2}$ so much?
\xe

By \citeauthor{chierchia2020}'s logic it follows that \textit{wh}-expressions
\textit{must} introduce \acp{dr} at their scope site, and this predicts that
\textit{wh}-scope should obviate \ac{wco}. This loses one of the core cases of
\ac{wco} -- \textit{wh}-movement can certainly \textit{not} obviate \ac{wco}.

\ex
\ljudge{*}Which boy$^{1}_{x}$ does his$_{1}$ father hate $t_{x}$?
\xe

\ex~
\ljudge{*}Which book$^{1}_{x}$ did John give none of it$_{1}$s fans $t_{x}$?
\xe



% \section{Rethinking the system}

% Conceptual issues:

% \begin{itemize}

%     \item Much like, e.g., \ac{dmg}, \citeauthor{chierchia2020}'s system
%     implicitly makes a syntactic distinction between two different kinds of
%     variables.\sidenote{See \citeauthor{barkerShan2008}'s
%     (\citeyear{barkerShan2008}) criticism of \ac{dmg} for related discussion.}
%     Can we do better? Continuations provide a way of doing scope-taking without
%     any need for traces, so this issue should (hopefully) dissolve if we shift
%     to a continuation-based theory.

% \end{itemize}

% \subsection{A more general dynamic system}

% \citet{charlow2019static} (see also \citealt{Charlowc}) generalizes orthodox
% dynamic semantics. We'll adopt his system here.

% \textit{Dynamic} $a$s are functions from an input assignment $g$, to a set of
% $a$s paired with output assignments $g'$.

% \ex
% $\type{D a} ≔ g → \set{(a,g)}$
% \xe

% \ex Unit (def.)\\
% $a^{η} ≔ λ g . \set{(a,g)}$
% \xe

% \todo[inline]{Redo definition of bind}

% \ex Bind (def.)\\
% $m^{⋆} ≔ λ k . λ g . \bigcup\limits_{(x,h) ∈ m g}\set{k x h}$\hfill$⋆: \type{D a → (a → D b) → D b}$
% \xe

% \ex
% $\ml{pro}_{n} ≔ λ g . \set{(g_{n},g)}$\hfill$\type{D e}$
% \xe

% \ex
% $\ml{some boy}_{n} ≔ λ g . \set{(x,g) | \ml{boy} x}$\hfill$\type{D e}$
% \xe

% \ex
% $m^{Δ_{n}} ≔ λ g . \bigcup\limits_{(x,h) ∈ m g}\set{(x,h^{[n → x]})}$\hfill$\type{D e → D e}$
% \xe

% \ex
% $\ml{Q} ≔ λ k . λ g . \set{(Q (λ x . (k x)^{↓_{g}})),g)}$\hfill$\ml{everyone}:\type{\semtower{D t}{e}}$
% \xe



\printbibliography


\end{document}
